---
title: "Patton College Students: Demographic and Socio-Economic Portrait"
output: 
  flexdashboard::flex_dashboard:
    theme: flatly
    social: ['facebook','twitter','linkedin']
    storyboard: true
    css: w3.css
runtime: shiny    
---

<script>

$(document).ready(function(){
    $("table").addClass("w3-table-all w3-hoverable");
});
</script>


```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(readxl)
library(leaflet)
library(htmltools)
library(plotly)
library(maps)
library(readxl)
library(ggplot2)

```


```{r global, include=FALSE}


students <- read_excel("geocoded_cleaned_final.xlsx")


```

### This map shows geographic distribution of Patton Colleges Students. The clusters represent number of students in the region.

```{r}

react_standing <- reactive({
  
if(input$standing =='All'){
  
  "Standing %in% c('Graduate','Senior', 'Freshman', 'Junior', 'Sophomore')"
  
} else if (input$standing =='Undergraduate') {
  
  "Standing %in% c('Senior', 'Freshman', 'Junior', 'Sophomore')"
  
} else{
  
  paste0("Standing == ",paste0("'",input$standing,"'"))
  
}

})


react_department <- reactive({

  if(input$department =="All"){

   "Dept.School %in% c('Educational Studies','Counseling & Higher Education', 'Teacher Education', 'Recreation and Sport Pedagogy', 'Human & Consumer Sciences','Education')"

  } else {

    paste0("Dept.School ==",paste0("'",input$department,"'"))
  }

})


temp <- reactive({
  filter_(students,react_standing(),react_department())

})


mypopup <- function(field1, field2, field3, field4){
  title <- '<b> STATS: </b>'
  
  income <- paste0('<tr> <td>',"Household Income:",'</td> <td>',paste0("$",as.character(prettyNum(field1,big.mark = ","))),'</td></tr>')
  college <- paste0('<tr> <td>',"College Degree: ", '</td> <td>',as.character(field2),"%",'</td></tr>')
  highschool <- paste0('<tr> <td>',"High School Degree: ",'</td> <td>' ,as.character(field3),"%",'</td>','</td></tr>')
  unemployment <- paste0('<tr> <td>',"Unemployment Rate: ", '</td> <td>', as.character(field4),"%",'</td>', '</td></tr>')
  paste0(title,"<table class='w3-table-all w3-hoverable'>", income, college,highschool,unemployment, '</table>')
}



human <- makeIcon(iconUrl ="student.png", iconRetinaUrl = "student.png",24,24)


renderLeaflet({

leaflet(temp()) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron", group="Default") %>% 
  addProviderTiles("Esri.WorldGrayCanvas", group="Esri") %>% 
  addMarkers(~lon ,~lat, clusterOptions = markerClusterOptions(),icon = human, popup = ~mypopup(Income, `College Degree Rate`, `High School Diploma Rate`, Unempl_rate)) %>% 
  addLayersControl(baseGroups=c("Default","Esri"),
  options = layersControlOptions(collapsed = FALSE)) 
  
  })


```

***

```{r}

selectInput("standing", label = "Standing:",
            choices = c("All","Undergraduate",unique(students$Standing)),selected = "All")

selectInput("department", label = "Department:",
            choices = c("All",unique(students$Dept.School)), selected = "All")

HTML("<p>Zoom in and click on the figure to get detailed information.</p>")


```



### This map is a snapshot of Patton College students. It shows geographic distribution and racial diversity of our students.

```{r}

pal <- colorFactor(c("red","red","green","black","magenta", "slateblue1"), domain = c( "Asian", "Black","HispLat"," Indian","Two or more","White"))

students <- students %>% filter(!is.na(Ethnicity))

temp1 <- reactive({
  
  
  if (input$race =="All"){
    
    students
    
  } else {
    
    students %>% filter(Ethnicity == input$race)
  }
  
  
})


renderLeaflet({

leaflet(temp1()) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron", group="Default") %>% 
  addProviderTiles("Esri.WorldGrayCanvas", group="Esri") %>% 
  addCircleMarkers(~lon ,~lat, radius = 4, color = ~pal(Ethnicity),stroke = FALSE, fillOpacity = 0.7,popup = ~mypopup(Income, `College Degree Rate`, `High School Diploma Rate`, Unempl_rate)) %>% 
  addLayersControl(baseGroups=c("Default","Esri"),
  options = layersControlOptions(collapsed = FALSE)) %>% 
  addLegend(pal = pal, values = ~ Ethnicity, opacity = 0.7)
  
})

```

***


```{r}



selectInput("race", label = "Race:",choices = c("All","Asian","Black","Indian","HispLat","Two or more","White"), selected = "All")


```

```{r}

knitr::kable(students %>% group_by(Ethnicity) %>% summarise(N=n()) %>%  mutate(Percent=round(N/sum(N)*100,1)) %>% rename(Race=Ethnicity))

```

### This map shows international students at Patton College both graduate and undergraduate.

```{r}

international <- read_excel("International Students.xlsx")
student <-international %>% select(Country, N) %>% filter(Country !="United States of America")

mapStates = map("world", student$Country, fill = TRUE, plot = FALSE)

renderLeaflet({
leaflet(mapStates) %>% addTiles() %>%
setView(lng =15 ,lat =15 , zoom = 2) %>%     
addProviderTiles("Esri.WorldGrayCanvas") %>%
addPolygons(stroke = F,smoothFactor = 0.2, color="red", fillOpacity = 0.3)

})

```


***

```{r}
student <- student %>% arrange(desc(N))
knitr::kable(student)

```


### This chart allows you to explore areas in which Patton College students live.

```{r}


mypal <- colorFactor(c("blue","green"), domain = c( "Selected", "Not Selected"))


container <- reactive({
  
  students[students$Income >= input$income[1] & students$Income <= input$income[2] & 
             students$`College Degree Rate`>= input$education[1] & students$`College Degree Rate` <= input$education[2],]
  
})


output$map1 <- renderLeaflet({
  
  leaflet(students) %>%
    addTiles() %>%
    addProviderTiles("CartoDB.Positron", group="Default") %>% 
    addProviderTiles("Esri.WorldGrayCanvas", group="Esri") %>% 
    fitBounds(lng1 = ~min(lon),lat1 =  ~min(lat),lng2 =  ~max(lon),lat2 =  ~max(lat)) %>% 
    
    addCircleMarkers(~lon ,~lat, radius = 4,color ="green", stroke = FALSE, fillOpacity = 0.5, popup = ~mypopup(Income, `College Degree Rate`, `High School Diploma Rate`, Unempl_rate)) %>% 
    #  addCircleMarkers(data = vlad(), lng= ~lon ,lat = ~lat, radius = 4,color ="green", stroke = FALSE, fillOpacity = 0.5, group = "Selected") %>% 
    
    addLayersControl(baseGroups=c("Default","Esri"),
                     #overlayGroups=c("Base","Selected"),                 
                     options = layersControlOptions(collapsed = FALSE))
  
})


observe({
  
  leafletProxy("map1", data = container()) %>% clearMarkers() %>% 

    addCircleMarkers(lng = ~lon ,lat = ~lat, radius = 4,color = "green", stroke = FALSE, fillOpacity = 0.5,popup = ~mypopup(Income, `College Degree Rate`, `High School Diploma Rate`, Unempl_rate))
})

leafletOutput("map1")   

```

***
```{r}

sliderInput(
  "income",
  h5("Median Household Income"),
  value = c(0, max(students$Income)),
  min = min(students$Income),
  max = max(students$Income)
  )
  
  sliderInput(
  "education",
  h5("Percent College Degree"),
  value = c(0, max(students$`College Degree Rate`)),
  min = min(students$`College Degree Rate`),
  max = max(students$`College Degree Rate`)
  )
  

```

